Attack Graph Generation for Microservice Architecture

### 2018-10-31 paper_own_docker/chapters/background_attack_graphs.tex
Feedback annotated with uncorrected corrections in yellow, notes with description are added and uploaded. Some very small corrections are undertaken in the paper as well.

Seyner et al. defined an attack graph as a tuple of states, transitions between the states, an initial state and success states. An initial state represents the state from which the attacker begins an attack and through a chain of atomic attacks attempts to reach one of the success states \cite{sheyner2002automated}. Ou et al. introduced the notion of a logical attack graph, which is a bipartite directed graph comprising fact and derivation nodes. Each fact node is labeled with a logical statement in the form of a predicate applied to its arguments, while each derivation node is labeled with an interaction rule used in the derivation step. The edges in a logical attack graph represent a "depends on" relation \cite{ou2006scalable}. Ingols et al. made a distinction between full, predictive and multiple-prerequisite (MP) attack graphs. A full graph is a directed acyclic graph comprising of nodes that represent hosts and edges that represent vulnerability instances. Predictive attack graphs use the same representation as full attack graphs with the only difference lying in the constraint of when the edges are added to the attack graph. Note that predictive graphs are generally smaller than full graphs. An MP attack is an attack graph with contentless edges, state nodes, vulnerability instance nodes and prerequisite nodes \cite{ingols2006practical}.

>>>>

Seyner et al. defined an attack graph as a tuple of states, transitions between the states, an initial state and success states. An initial state represents the state from which the attacker begins an attack and through a chain of atomic attacks attempts to reach one of the success states \cite{sheyner2002automated}. Ou et al. introduced the notion of a logical attack graph, which is a bipartite directed graph comprising fact and derivation nodes. Each fact node is labeled with a logical statement in the form of a predicate applied to its arguments, while each derivation node is labeled with an interaction rule used in the derivation step. The edges in a logical attack graph represent a "depends on" relation \cite{ou2006scalable}. Ingols et al. made a distinction between full, predictive, and multiple-prerequisite (MP) attack graphs. A full graph is a directed acyclic graph comprising nodes that represent hosts and edges that represent vulnerability instances. Predictive attack graphs use the same representation as full attack graphs, with the only difference lying in the constraint of when the edges are added to the attack graph. Note that predictive graphs are generally smaller than full graphs. An MP attack is an attack graph with contentless edges, state nodes, vulnerability instance nodes, and prerequisite nodes \cite{ingols2006practical}.

### 2018-10-31 paper_own_docker/chapters/evaluation.tex
Feedback annotated with uncorrected corrections in yellow, notes with description are added and uploaded. Some very small corrections are undertaken in the paper as well.

@@

Real-world microservice systems comprise many containers that run different technologies with various degrees of connectivity among each other. This raises the need for a robust and scalable attack graph generator. First, we demonstrate use cases in Subsection \ref{chap:heterogenious_systems}. We then examine how others have evaluated their systems. In Subsection \ref{chap:scalability_eval}, we discuss experiments conducted to test the scalability of the proposed system with different numbers of containers and varying degrees of connectivity. Note that all experiments were performed on an Intel(R) Core(TM) i5-7200U CPU (2.50GHz) with 8 GB of RAM running Ubuntu 16.04.3 LTS.

Modern microservice architectures use many different technologies,  different numbers of containers, various degrees of connectivity, and have different numbers of vulnerabilities. Therefore it is critically important to demonstrate that an attack graph generator works well in such heterogeneous scenarios. Here, we tested the proposed system on real and slightly modified GitHub examples (Table \ref{table_technologies}). We employed test examples that are publicly available to facilitate potential future comparison characterized by different system properties (e.g., topologies, technologies and vulnerabilities) and different usage domains. We also had to consider the fact that an overwhelming majority of publicly-available examples are small, i.e., only one or a few containers, which made finding appropriate test examples challenging. The resulting examples are as follows: NetflixOSS, the Atsea Sample Shop App, and the JavaEE demo. NetflixOSS is a microservice system provided by Netflix comprising of 10 containers and uses many tools, e.g., Spring Cloud, Netflix Ribbon, and Netflix Eureka. The Atsea Sample Shop App is an e-commerce sample web application comprising four containers and uses Spring Boot, React, NGINX and PostgreSQL. The JavaEE demo is a sample application for browsing movies comprising only two containers and uses JavaEE, React and Tomcat EE. We ran the attack graph generator and manually verified the resulting attack graphs for the small examples based on domain knowledge under the assumption that the output from Clair, the NVD attack vectors \cite{booth2013national}, and the preconditions and postconditions from Aksu et al. \cite{aksu2018automated} are \textit{correct}. After running the proposed attack graph generator, the attack graphs for the Atsea Sample Shop App and JavaEE demo were small as expected, containing only a few nodes and edges. The structure of the NetflixOSS attack graph demonstrated a nearly linear structure in which each node was connected to a small number of other nodes to form a chain of attacks. This linearity is due to the fact that each container is connected to only a few other containers to reduce unnecessary communication and increase encapsulation. Therefore, based on this degree of connectivity an attacker needs to perform multiple intermediate steps to reach the target container. Note that all examples terminated, there were no directed edges from containers with higher privileges to lower privileges, no duplication of nodes, and no reflexive edges were observed, which is in line with the monotonicity property. In addition, the run time of the proposed system with each example was short, however, additional scalability tests were required. Therefore, the Phpmailer and Samba system was extended and employed as an artificial example to perform scalability tests. This is discussed in the following subsection.

@@

Sheyner et al. \cite{sheyner2002automated} tested their system (NuSMV) using both small and extended examples. The attack graph in the larger example has 5948 nodes and 68364 edges. The time required for NuSMV to execute this configuration was two hours; however, the model checking component took four minutes. The authors claim that the performance bottleneck resides in the graph generation procedure. Ingols et al. \cite{ingols2006practical} tested their system on a network of 250 hosts. They continued the study on a simulated network with 50000 hosts in under four minutes. Although their method yields better performance than NuSMV, their evaluation was based on a MP graph, which differs from our target graph. Ou et al. \cite{ou2006scalable} provided a more extensive study wherein they tested their system (MulVAL) using more examples. They state that the asymptotic CPU time was between $O(n^2)$ and $O(n^3)$, where n is the number of nodes (hosts). With 1000 fully-connected nodes, their system required more than 1000 seconds to execute. % The authors also provide an evaluation where he MulVAL clearly outperforms the Sheyner's system.

@@

In summary, we found that the proposed algorithm generates attack graphs efficiently, i.e., it handles a system with 1000 containers in 13 minutes. Considering the strongly-connected system employed in the experiment and the high number of vulnerabilities in it, we consider that the results demonstrate that the proposed system is a practical solution that can be used as part of the continuous delivery processes of real-world systems.

>>>>

Real-world microservice systems comprise many containers that run different technologies with various degrees of connectivity among each other. This raises the need for a robust and scalable attack graph generator. We demonstrate use cases in Subsection \ref{chap:heterogenious_systems}. We then examine how others have evaluated their systems. In Subsection \ref{chap:scalability_eval}, we discuss experiments conducted to test the scalability of the proposed system with different numbers of containers and varying degrees of connectivity. Note that all experiments were performed on an Intel(R) Core(TM) i5-7200U CPU (2.50GHz) with 8 GB of RAM running Ubuntu 16.04.3 LTS.

@@

Modern microservice architectures use many different technologies,  different numbers of containers, various degrees of connectivity, and have different numbers of vulnerabilities. Therefore, it is critically important to demonstrate that an attack graph generator works well in such heterogeneous scenarios. Here, we tested the proposed system on real and slightly modified GitHub examples (Table \ref{table_technologies}). We employed test examples that are publicly available to facilitate potential future comparison characterized by different system properties (e.g., topologies, technologies and vulnerabilities) and different usage domains. We also had to consider the fact that an overwhelming majority of publicly-available examples are small, i.e., only one or a few containers, which made finding appropriate test examples challenging. The resulting examples are as follows: NetflixOSS, the Atsea Sample Shop App, and the JavaEE demo. NetflixOSS is a microservice system provided by Netflix comprising 10 containers and uses many tools, e.g., Spring Cloud, Netflix Ribbon, and Netflix Eureka. The Atsea Sample Shop App is an e-commerce sample web application comprising four containers and uses Spring Boot, React, NGINX, and PostgreSQL. The JavaEE demo is a sample application for browsing movies comprising only two containers and uses JavaEE, React, and Tomcat EE. We ran the attack graph generator and manually verified the resulting attack graphs for the small examples based on domain knowledge under the assumption that the output from Clair, the NVD attack vectors \cite{booth2013national}, and the preconditions and postconditions from Aksu et al. \cite{aksu2018automated} are \textit{correct}. After running the proposed attack graph generator, the attack graphs for the Atsea Sample Shop App and JavaEE demo were small as expected, containing only a few nodes and edges. The structure of the NetflixOSS attack graph demonstrated a nearly linear structure in which each node was connected to a small number of other nodes to form a chain of attacks. This linearity is due to the fact that each container is connected to only a few other containers to reduce unnecessary communication and increase encapsulation. Therefore, based on this degree of connectivity, an attacker needs to perform multiple intermediate steps to reach the target container. Note that all examples terminated, there were no directed edges from containers with higher privileges to lower privileges, no duplication of nodes, and no reflexive edges were observed, which is in line with the monotonicity property. In addition, the run time of the proposed system with each example was short, however, additional scalability tests were required. Therefore, the Phpmailer and Samba system was extended and employed as an artificial example to perform scalability tests. This is discussed in the following subsection.

@@

Sheyner et al. \cite{sheyner2002automated} tested their system (NuSMV) using both small and extended examples. The attack graph in the larger example has 5948 nodes and 68364 edges. The time required for NuSMV to execute this configuration was two hours; however, the model checking component took four minutes. The authors claim that the performance bottleneck resides in the graph generation procedure. Ingols et al. \cite{ingols2006practical} tested their system on a network of 250 hosts. They continued the study on a simulated network with 50000 hosts in under four minutes. Although their method yields better performance than NuSMV, their evaluation was based on a MP graph, which differs from our target graph. Ou et al. \cite{ou2006scalable} provided a more extensive study, wherein they tested their system (MulVAL) using more examples. They state that the asymptotic CPU time was between $O(n^2)$ and $O(n^3)$, where n is the number of nodes (hosts). With 1000 fully-connected nodes, their system required more than 1000 seconds to execute. % The authors also provide an evaluation where he MulVAL clearly outperforms the Sheyner's system.

@@

In summary, we found that the proposed algorithm generates attack graphs efficiently, i.e., it handles a system with 1000 containers in 13 minutes. Considering the strongly-connected system employed in the experiment and the high number of vulnerabilities in this system, we consider that the results demonstrate that the proposed system is a practical solution that can be used as part of the continuous delivery processes of real-world systems.

### 2018-10-31 paper_own_docker/chapters/method.tex
Feedback annotated with uncorrected corrections in yellow, notes with description are added and uploaded. Some very small corrections are undertaken in the paper as well.

\section{METHOD}
\label{chap:method}
In Subsection \ref{chap:mapping}, we discuss how the existing components of attack graph generation for a computer network are mapped to a microservice environment, and the concepts are illustrated using a small example. Then, in Subsection \ref{chap:technical}, we present the tools we use to achieve this mapping and provide an overview of the proposed system and its components, i.e., the Topology Parser (Subsection \ref{chap:topology_p}), the Vulnerability Parser (Subsection \ref{chap:vulnerability_p}) and the Attack Graph Generator (Subsection \ref{chap:attack_graph_p}). In addition, the Breath-first Search (BFS) graph traversal algorithm is discussed in Subsection \ref{chap:bfs}.

@@

Here, we present a small example to demonstrate how attack graph generation works in practice. The example is taken from the Netflix OSS GitHub repository. The Netflix OSS example is a Spring Cloud-based microservice architecture that uses the following microservices: Service Discovery (Eureka), Circuit Breaker (Hystrix), Intelligent Routing (Zuul), and Client Side Load Balancing (Ribbon). Figure \ref{TopologyGraph} shows a subset of the example topology where each node denotes a container and each edge is a connection between two containers if one calls the other. The topology comprises an "Outside" node and a "Docker daemon" node, as well as Zuul, Eureka and other nodes. According to Netflix, Zuul is an edge service that provides dynamic routing, monitoring, resilience, and security functionalities. Eureka is a Representational State Transfer (REST) based service primarily used in the cloud for locating services for load balancing and fail-over of middle-tier servers. Figure \ref{AttackGraph} shows a part of the corresponding attack graph, where a node is a pair of the image and its privilege, while an edge represents an atomic attack. Parts of both graphs have been omitted intentionally for simplicity. An example path an attacker would take could be to first attack the Zuul container by exploiting the CVE-2016-10249 vulnerability by crafting an image file, which triggers a heap-based buffer overflow\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2016-10249}} and gains the USER privilege.  With this USER privilege, an attacker can exploit the CVE-2015-7554 vulnerability on the same container via crafted field data in an extension tag in a TIFF image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2015-7554}} to gain the ADMIN privilege. Once the ADMIN privilege has been obtained on the Zuul container, the attacker can attack the Eureka container by exploiting CVE-2017-7600 via another crafted image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2017-7600}} and gain the ADMIN privilege. Note that this is not the only path the attacker can take to obtain ADMIN privileges on the Eureka container. Another path would be to exploit the CVE-2018-1124 vulnerability by creating entries in the file system (procfs) by starting processes, which could result in crashes or arbitrary code execution\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2018-1124}}. This vulnerability can be exploited by having only the USER privilege on Zuul to gain the ADMIN privileges of the Eureka container directly. Our attack graph generator shows both paths because it is of interest to identify all possible routes through which a container can be compromised.

@@

In the preprocessing step, we use Clair to generate the vulnerabilities for a given image. Clair is a vulnerability scanner that inspects a Docker image and generates its vulnerabilities by providing a \textit{CVE-ID}, a description and an attack vector for each vulnerability. An attack vector is an entity that describes which conditions and effects are connected to the given vulnerability. We collect the fields in the attack vector as defined by the NVD \cite{booth2013national}, i.e.,Access Vector (Local, Adjacent Network, Network), Access Complexity (Low, Medium, High), Authentication (None, Single, Multiple), Confidentiality Impact (None, Partial, Complete), Integrity Impact (None, Partial, Complete), and Availability Impact (None, Partial, Complete). Since Clair does not provide a command line interface to analyze a Docker image, we use Clairctl to analyze a complete Docker image.

@@

After preprocessing, the vulnerabilities are parsed and their preconditions and postconditions are extracted. Together with the topology, they are fed into a BFS algorithm. BFS is a popular search algorithm that traverses a graph by first looking at the neighbors of a given node before diving deeper into the graph. The pseudocode for our modified BFS algorithm is given in Algorithm \ref{BFSalgorithm}. This algorithm requires a topology, a dictionary of the exploitable vulnerabilities and a list of nodes with privileged access as input. The output comprises nodes and edges that form the attack graph. In Algorithm \ref{BFSalgorithm}, "topology" (Subsection \ref{chap:topology_p}) provides information about the connectivity between containers, "cont\_expl" (Subsections \ref{chap:vulnerability_p} and \ref{chap:attack_graph_p}) contains information about which vulnerabilities can be attacked (with their preconditions and postconditions), and "priv\_acc" (Subsection \ref{chap:topology_p}) is an array of nodes with high (i.e., ADMIN) permissions to the Docker daemon. First, the algorithm initializes the nodes, edges, queue, and passed nodes (lines \ref{alg:init_1} and \ref{alg:init_2}). Then, it generates the attacker node (line \ref{alg:attackNodeInit}) as the node where the attack begins. The attacker node is a combination of the image name ("outside") and the privilege level (ADMIN). Then, in a while loop (line \ref{alg:while}), the algorithm iterates through each node (line \ref{alg:get_node}), checks the given node's neighbors (line \ref{alg:neigh}), and adds the edges if the conditions are satisfied (lines \ref{alg:daemonEdge}, \ref{alg:privEdge} and \ref{alg:addEdge}). If a neighbor was not passed, then it is added to the queue (line \ref{alg:putQueue}). The algorithm terminates when the queue is empty (line \ref{alg:while}). Furthermore, BFS is characterized by the following properties.

>>>>

\section{METHOD}
\label{chap:method}
In Subsection \ref{chap:mapping}, we discuss how the existing components of attack graph generation for a computer network are mapped to a microservice environment, and the concepts are illustrated using a small example. Then, in Subsection \ref{chap:technical}, we present the tools we use to achieve this mapping and provide an overview of the proposed system and its components, i.e., the Topology Parser (Subsection \ref{chap:topology_p}), the Vulnerability Parser (Subsection \ref{chap:vulnerability_p}), and the Attack Graph Generator (Subsection \ref{chap:attack_graph_p}). In addition, the Breath-first Search (BFS) graph traversal algorithm is discussed in Subsection \ref{chap:bfs}.

@@

Here, we present a small example to demonstrate how attack graph generation works in practice. The example is taken from the Netflix OSS GitHub repository. The Netflix OSS example is a Spring Cloud-based microservice architecture that uses the following microservices: Service Discovery (Eureka), Circuit Breaker (Hystrix), Intelligent Routing (Zuul), and Client Side Load Balancing (Ribbon). Figure \ref{TopologyGraph} shows a subset of the example topology, where each node denotes a container and each edge is a connection between two containers if one calls the other. The topology comprises an "Outside" node and a "Docker daemon" node, as well as Zuul, Eureka, and other nodes. According to Netflix, Zuul is an edge service that provides dynamic routing, monitoring, resilience, and security functionalities. Eureka is a Representational State Transfer (REST) based service primarily used in the cloud for locating services for load balancing and fail-over of middle-tier servers. Figure \ref{AttackGraph} shows a part of the corresponding attack graph, where a node is a pair of the image and its privilege, while an edge represents an atomic attack. Parts of both graphs have been omitted intentionally for simplicity. An example path an attacker would take could be to first attack the Zuul container by exploiting the CVE-2016-10249 vulnerability by crafting an image file, which triggers a heap-based buffer overflow\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2016-10249}} and gains the USER privilege.  With this USER privilege, an attacker can exploit the CVE-2015-7554 vulnerability on the same container via crafted field data in an extension tag in a TIFF image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2015-7554}} to gain the ADMIN privilege. Once the ADMIN privilege has been obtained on the Zuul container, the attacker can attack the Eureka container by exploiting CVE-2017-7600 via another crafted image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2017-7600}} and gain the ADMIN privilege. Note that this is not the only path the attacker can take to obtain ADMIN privileges on the Eureka container. Another path would be to exploit the CVE-2018-1124 vulnerability by creating entries in the file system (procfs) by starting processes, which could result in crashes or arbitrary code execution\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2018-1124}}. This vulnerability can be exploited by having only the USER privilege on Zuul to gain the ADMIN privileges of the Eureka container directly. Our attack graph generator shows both paths because it is of interest to identify all possible routes through which a container can be compromised.

@@

In the preprocessing step, we use Clair to generate the vulnerabilities for a given image. Clair is a vulnerability scanner that inspects a Docker image and generates its vulnerabilities by providing a \textit{CVE-ID}, a description and an attack vector for each vulnerability. An attack vector is an entity that describes which conditions and effects are connected to the given vulnerability. We collect the fields in the attack vector as defined by the NVD \cite{booth2013national}, i.e., Access Vector (Local, Adjacent Network, Network), Access Complexity (Low, Medium, High), Authentication (None, Single, Multiple), Confidentiality Impact (None, Partial, Complete), Integrity Impact (None, Partial, Complete), and Availability Impact (None, Partial, Complete). Since Clair does not provide a command line interface to analyze a Docker image, we use Clairctl to analyze a complete Docker image.

@@

After preprocessing, the vulnerabilities are parsed and their preconditions and postconditions are extracted. Together with the topology, they are fed into a BFS algorithm. BFS is a popular search algorithm that traverses a graph by first looking at the neighbors of a given node before diving deeper into the graph. The pseudocode for our modified BFS algorithm is given in Algorithm \ref{BFSalgorithm}. This algorithm requires a topology, a dictionary of the exploitable vulnerabilities, and a list of nodes with privileged access as input. The output comprises nodes and edges that form the attack graph. In Algorithm \ref{BFSalgorithm}, "topology" (Subsection \ref{chap:topology_p}) provides information about the connectivity between containers, "cont\_expl" (Subsections \ref{chap:vulnerability_p} and \ref{chap:attack_graph_p}) contains information about which vulnerabilities can be attacked (with their preconditions and postconditions), and "priv\_acc" (Subsection \ref{chap:topology_p}) is an array of nodes with high (i.e., ADMIN) permissions to the Docker daemon. First, the algorithm initializes the nodes, edges, queue, and passed nodes (lines \ref{alg:init_1} and \ref{alg:init_2}). Then, it generates the attacker node (line \ref{alg:attackNodeInit}) as the node where the attack begins. The attacker node is a combination of the image name ("outside") and the privilege level (ADMIN). Then, in a while loop (line \ref{alg:while}), the algorithm iterates through each node (line \ref{alg:get_node}), checks the given node's neighbors (line \ref{alg:neigh}), and adds the edges if the conditions are satisfied (lines \ref{alg:daemonEdge}, \ref{alg:privEdge} and \ref{alg:addEdge}). If a neighbor was not passed, then it is added to the queue (line \ref{alg:putQueue}). The algorithm terminates when the queue is empty (line \ref{alg:while}). Furthermore, BFS is characterized by the following properties.
